---
title: "Exploratory Analysis of Sierra Amphibians"
description: |
  An exploration of amphibian abundance from the Sierra Lakes Inventory Project.
author:
  - name: Halley McVeigh
    url: {}
date: 2022-02-21
output:
  distill::distill_article:
    self_contained: false
    code_folding: hide
    theme: sandstone
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(readxl)
library(here)
library(lubridate)
library(patchwork)
```

## This project involves wrangling data for Sierra amphibians and creating a compound data visualization figure. It is an exploration of amphibian abundance data recorded by the Sierra Lakes Inventory Project (SLIP).

### Data citation: *Knapp,R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative.* https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3




```{r}
# Read in amphibian data and wrangle

amphibians <- readxl::read_xlsx(here("data", "sierra_amphibians.xlsx"))

rana_muscosa <- amphibians %>% 
  filter(amphibian_species == "RAMU") %>% #filter to include only mountain yellow-legged frog
  filter(amphibian_life_stage == c("SubAdult", "Tadpole", "Adult")) %>% 
  mutate(date = ymd(survey_date)) #convert date to date format

rana_muscosa$date <- as.numeric(format(rana_muscosa$date, "%Y"))  #extract just the year
  
rana_muscosa_summary <- rana_muscosa %>% 
  group_by(date, amphibian_life_stage) %>%  #group by year and life stage
  summarize(frog_count = sum(amphibian_number, na.rm = TRUE)) #and summarize counts of frogs
```

### Graph #1: Plot of adult, subadult, and tadpole mountain yellow-legged frog counts by life stage
```{r}
#create ggplot of counts, facet wrapped by life stage
graph1 <- ggplot(data = rana_muscosa_summary, aes(x = date, y = frog_count, color = amphibian_life_stage)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(title = "Mountain Yellow-Legged Frog Counts from 1995 to 2002",
       x = "Year",
       y= "Count",
       color = "Life Stage") +
  facet_wrap(~amphibian_life_stage) +
  theme(legend.position = "none")

graph1
```

### Graph #2: Plot combining adult and subadult mountain yellow-legged frog counts observed in 5 lakes with greatest total observed counts

```{r}

amphibians$lake_id <-as.character(as.numeric(amphibians$lake_id))

rana_muscosa_top5 <- amphibians %>% 
  filter(amphibian_life_stage %in% c("SubAdult", "Adult")) %>% #filter to include only adult and subadult frogs (no tadpoles)
  filter(amphibian_location == "in lake") %>% #only include frogs observed in the lake
  filter(amphibian_species == "RAMU") %>% #filter for rana muscosa species
  group_by(lake_id) %>% 
  summarize(rana_muscosa_count = sum(amphibian_number)) %>%  #total frog counts per each lake
  mutate(lake_id = fct_reorder(lake_id, rana_muscosa_count)) %>% #arrange in descending order
  slice_max(rana_muscosa_count, n=5) #keep top 5 highest values of frog counts per lake

rana_muscosa_top5$lake_id = paste('Lake', rana_muscosa_top5$lake_id, sep = ' ')#%>%
  #mutate(lake_id = fct_reorder(lake_id, rana_muscosa_count))
  
# create ggplot
graph2 <- ggplot(data= rana_muscosa_top5, aes(x=reorder(lake_id, -rana_muscosa_count), y=rana_muscosa_count)) +
  geom_col(fill = "cyan4", color = "black") +
  labs(title = "Top 5 lakes with the greatest counts of \nMountain yellow-legged adult and subadult frogs", 
       x= "Lake ID",
       y = "Count (adults and subadults)") +
  theme_minimal() +
  coord_flip()

graph2
```

### Combine both plots into a compound figure using `patchwork` package

```{r}

graph1 + graph2 + plot_layout(ncol = 1, heights = c(2,2)) #combining graph 1 and graph 2 on the same figure
```

**Figure 3.** Top: The change in Mountain Yellow-Legged Frog counts from 1995 to 2002 for different life stages - adults, subadults and tadpoles. Bottom: The top 5 lakes with the highest counts of adult and subadult Mountain Yellow-Legged Frogs.



